---
title: "DATA REPROCESSING"
author: "THEDION V. DIAM JR."
date: "2022-12-16"
output: pdf_document
---

# LOAD PACKAGES

```{r}
# Helper packages
library(readr)    # loading dataset
library(dplyr)    # for data wrangling
library(ggplot2)  # for awesome graphics
library(tidyverse)# for filtering 
library(rsample)  # for creating validation splits
library(bestNormalize) # for normalizing the dataset
library(stringr)       # for string functionality
library(gridExtra)     # for manipulaiting the grid

# Modeling packages
library(cluster)       # for general clustering algorithms
library(factoextra)    # for visualizing cluster results
library(ranger)     # a c++ implementation of random forest 
library(h2o)        # a java-based implementation of random forest
h2o.init()
```

# Load Data Sets
The data contains 197 rows and 431 columns with *Failure.binary* binary output.
```{r}
library(readr)
rawd <- read_csv("D:/DIAM/FP-DATA.csv")
```


#=================================================== Reprocessing the Raw Data ===================================================#
```{r}
library(tidyverse)
library(bestNormalize)
```

# Check for null and missing values
Using *anyNA()* function, We can determine if any missing values in our data.
```{r}
anyNA(rawd)

#The result shows either *True* or *False*. If True, omit the missing values using *na.omit()*
  
#[1] FALSE

#Thus, our data has no missing values.
```

# Check for Normality of the Data
We used *Shapiro-Wilk's Test* to check the normality of the data.

```{r,warning=F}
rd <- rawd%>%select_if(is.numeric) 
rd <- rd[,-1]
test <- apply(rd,2,function(x){shapiro.test(x)})
```


To have the list of p-value of all variables, the *unlist()* function is used and convert a list to vector.
```{r}
pvalue_list <- unlist(lapply(test, function(x) x$p.value))
```


```{r}
sum(pvalue_list<0.05)  # not normally distributed
sum(pvalue_list>0.05)  # normally distributed
test$Entropy_cooc.W.ADC

# [1] 428
# [1] 1

#  Thus, we have 428 variables that are not normally distributed and Entropy_cooc.W.ADC is normally distributed.
```

We use *orderNorm()* function, the *x.t*	is the elements of orderNorm() function transformed original data.Using the *Shapiro-Wilk's Test*
```{r,warning=F}
TRDrawd=rawd[,c(3,5:length(names(rawd)))]

TRDrawd=apply(TRDrawd,2,orderNorm)
TRDrawd=lapply(TRDrawd, function(x) x$x.t)
TRDrawd=TRDrawd%>%as.data.frame()
test=apply(TRDrawd,2,shapiro.test)
test=unlist(lapply(test, function(x) x$p.value))
```

#Testing Data 
```{r,warning=F}
sum(test <0.05)  # not normally distributed
sum(test >0.05)  # normally distributed

#[1] 0
#[1] 428

# Thus, our data is normally distributed.
```


```{r}
rawd[,c(3,5:length(names(rawd)))]=TRDrawd
```

Get the correlation of the whole data expect the categorical variables
```{r}
CorMatrix=cor(rawd[,-c(1,2)])
heatmap(CorMatrix,Rowv=NA,Colv=NA,scale="none",revC = T)
```

#Splitting the Data
Split the data into training (80%) and testing (20%).
```{r}
rawd$Institution=as.factor(rawd$Institution)
rawd$Failure.binary=as.factor(rawd$Failure.binary)
```

```{r}
splitter <- sample(1:nrow(rawd), round(nrow(rawd) * 0.8))
trainND <- rawd[splitter, ]
testND  <- rawd[-splitter, ]
```

The data frame output of data reprocessing will be converted into to "csv", which will be used for entire project.

# Load new Data 
```{r}
Final<-  read_csv("D:/DIAM/newdat.csv")
View(Final)

