---
title: "STAT 325 Model 1"
author: "THEDION V. DIAM JR."
date: "2022-12-14"
output: html_document
---
# Load Data Sets
The data contains 197 rows and 431 columns with *Failure.binary* binary output.
```{r}
rawd <- read.csv("C:/Users/redee/OneDrive/Desktop/STAT 325 Final Project/FP DATA.csv")
```


#=================================================== Reprocessing the Raw Data ===================================================#
```{r}
library(tidyverse)
library(bestNormalize)
```

# Check for null and missing values
Using *anyNA()* function, We can determine if any missing values in our data.
```{r}
anyNA(rawd)

#The result shows either *True* or *False*. If True, omit the missing values using *na.omit()*
  
#[1] FALSE

#Thus, our data has no missing values.
```

# Check for Normality of the Data
We used *Shapiro-Wilk's Test* to check the normality of the data.

```{r,warning=F}
rd <- rawd%>%select_if(is.numeric) 
rd <- rd[,-1]
test <- apply(rd,2,function(x){shapiro.test(x)})
```


To have the list of p-value of all variables, the *unlist()* function is used and convert a list to vector.
```{r}
pvalue_list <- unlist(lapply(test, function(x) x$p.value))
```


```{r}
sum(pvalue_list<0.05)  # not normally distributed
sum(pvalue_list>0.05)  # normally distributed
test$Entropy_cooc.W.ADC

# [1] 428
# [1] 1

#  Thus, we have 428 variables that are not normally distributed and Entropy_cooc.W.ADC is normally distributed.
```

We use *orderNorm()* function, the *x.t*	is the elements of orderNorm() function transformed original data.Using the *Shapiro-Wilk's Test*
```{r,warning=F}
TRDrawd=rawd[,c(3,5:length(names(rawd)))]

TRDrawd=apply(TRDrawd,2,orderNorm)
TRDrawd=lapply(TRDrawd, function(x) x$x.t)
TRDrawd=TRDrawd%>%as.data.frame()
test=apply(TRDrawd,2,shapiro.test)
test=unlist(lapply(test, function(x) x$p.value))
```

#Testing Data 
```{r,warning=F}
sum(test <0.05)  # not normally distributed
sum(test >0.05)  # normally distributed

#[1] 0
#[1] 428

# Thus, our data is normally distributed.
```


```{r}
rawd[,c(3,5:length(names(rawd)))]=TRDrawd
```

Get the correlation of the whole data expect the categorical variables
```{r}
CorMatrix=cor(rawd[,-c(1,2)])
heatmap(CorMatrix,Rowv=NA,Colv=NA,scale="none",revC = T)
```

#Splitting the Data
Split the data into training (80%) and testing (20%).
```{r}
rawd$Institution=as.factor(rawd$Institution)
rawd$Failure.binary=as.factor(rawd$Failure.binary)
```

```{r}
splitter <- sample(1:nrow(rawd), round(nrow(rawd) * 0.8))
trainND <- rawd[splitter, ]
testND  <- rawd[-splitter, ]
```

The data frame output of data reprocessing will be converted into to "csv", which will be used for entire project.

# Load new Data 
```{r}
Final<- read.csv("C:/Users/redee/OneDrive/Desktop/STAT 325 Final Project/newdat.csv")
View(Final)
```

#=============================================GRADIENT BOOSTING===================================================================#

```{r}
# Helper packages
library(dplyr)    # for general data wrangling needs
library(tidyverse)
library(rsample)
library(ROCR)
library(pROC)
# Modeling packages
library(gbm)      # for original implementation of regular and stochastic GBMs
library(h2o)      # for a java-based implementation of GBM variants
library(xgboost)  # for fitting extreme gradient boosting
```

```{r}
set.seed(123)  # for reproducibility



Final$Institution=as.factor(Final$Institution)
split <- initial_split(Final, strata = "Failure.binary")
trainFinal <- training(split)
testFinal <- testing(split)
```

```{r}
gradientBoostingModel_1 <- gbm(
  formula = Failure.binary ~ .,
  data = trainFinal,
  distribution = "bernoulli",  # SSE loss function
  n.trees = 1000,
  shrinkage = 0.1,
  n.minobsinnode = 10,
  cv.folds = 10

)
```


```{r}
# find index for number trees with minimum CV error
best <- which.min(gradientBoostingModel_1$cv.error)
```

```{r}
# get MSE and compute RMSE
sqrt(gradientBoostingModel_1$cv.error[best])
```
```{r}
# plot error curve
gbm.perf(gradientBoostingModel_1, method = "cv")
```
```{r}
# create grid search
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  logloss = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {
  
  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = Failure.binary ~ .,
      data = trainFinal,
      distribution = "bernoulli",
      n.trees = 1000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
    )
  })
  
  # add SSE, trees, and training time to results
  hyper_grid$logloss[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]
  
}
```

```{r}
# results
arrange(hyper_grid, logloss)
```


```{r}
# search grid
hyper_grid <- expand.grid(
  n.trees = 1000,
  shrinkage = 0.01,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)

)
```

```{r}
# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Failure.binary ~ .,
    data = trainFinal,
    distribution = "bernoulli",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))

}
```

```{r}
# perform search grid with functional programming
hyper_grid$logloss <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
  )
)

# results
arrange(hyper_grid, logloss)
```
```{r}
# refined hyperparameter grid
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "logloss",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)
```

```{r}
# perform grid search 
trainFinal$Failure.binary=as.factor(trainFinal$Failure.binary)
h2o.shutdown()
h2o.init()
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  y = "Failure.binary",
  training_frame = as.h2o(trainFinal),
  hyper_params = hyper_grid,
  ntrees = 10,#supposedly 6000
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 5,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  stopping_metric="logloss",
  search_criteria = search_criteria,
  seed = 123

)
```

```{r}
# collect the results and sort by our model performance metric of choice
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "logloss", 
  decreasing = FALSE
)
```

```{r}
grid_perf
```
```{r}
# Grab the model_id for the top model, chosen by cross validation error
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let’s get performance metrics on the best model
h2o.performance(model = best_model, xval = TRUE)
```
```{r}
library(recipes)
xgb_prep <- recipe(Failure.binary ~ ., data = trainFinal) %>%
  step_integer(all_nominal()) %>%
  prep(training = trainFinal, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Failure.binary")])
Y <- xgb_prep$Failure.binary
Y=as.numeric(Y)-1
```

```{r}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 1000,
  objective = "binary:logistic",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)

```

```{r}
# minimum test CV RMSE
min(ames_xgb$evaluation_log$test_logloss_mean)
```

```{r}
# hyperparameter grid
hyper_grid <- expand.grid(
  eta = 0.01,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.5, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  logloss = 0,          # a place to dump RMSE results
  trees = 0          # a place to dump required number of trees
)

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 1000,#supposedly 4000
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i], 
      lambda = hyper_grid$lambda[i], 
      alpha = hyper_grid$alpha[i]
    ) 
  )
  hyper_grid$logloss[i] <- min(m$evaluation_log$test_logloss_mean)
  hyper_grid$trees[i] <- m$best_iteration
}
```

```{r}
# results
hyper_grid %>%
  filter(logloss > 0) %>%
  arrange(logloss) %>%
  glimpse()
```

```{r}
# optimal parameter list
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)
```

```{r}
# 
# # train final model
# trainFinal$Institution=fct_recode(trainFinal$Institution, "1" = "A", "2" ="B","3"="C","4"="D")
# trainFinal$Institution=as.numeric(trainFinal$Institution)
# trainFinal=as.matrix(trainFinal)

xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 3944,
  objective = "binary:logistic",
  verbose = 0
)
```

```{r}
# Compute predicted probabilities on training data
m1_prob <- predict(xgb.fit.final, X, type = "prob")

# Compute AUC metrics for cv_model1,2 and 3 
perf1 <- prediction(m1_prob,trainFinal$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")


# Plot ROC curves for cv_model1,2 and 3 
plot(perf1, col = "black", lty = 2)


# ROC plot for training data
roc( trainFinal$Failure.binary ~ m1_prob, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)

xgb_prep <- recipe(Failure.binary ~ ., data = testFinal) %>%
  step_integer(all_nominal()) %>%
  prep(training = testFinal, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Failure.binary")])

# Compute predicted probabilities on training data
m2_prob <- predict(xgb.fit.final, X, type = "prob")

# Compute AUC metrics for cv_model1,2 and 3 
perf2 <- prediction(m2_prob,testFinal$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")


# Plot ROC curves for cv_model1,2 and 3 
plot(perf2, col = "black", lty = 2)


# ROC plot for training data
roc( testFinal$Failure.binary ~ m2_prob, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)

```

```{r}
# variable importance plot
vip::vip(xgb.fit.final,num_features=20) 
```